Parallel Multi-objective Optimization Method for Finding Complete Set of Weakly Efficient Solutions

### 2017-10-09 2017/URAL-PDC/u_pdc_paper.tex
u-pdc: fix some issues with English language

\begin{abstract}
This work considers a parallel algorithm for the multi-objective optimization. The
algorithm is based on the application of the information-statistical approach for solving
reduced scalar problem, the set of the global optima in which coincides to the set of the
weakly efficient solutions of the initial multi-objective problem. In the present work the parallelization scheme, which is common for all information-statistical global search algorithms,
was applied to the sequential algorithm of the multi-objective optimization. Also, in this
work one of the techniques of accounting for the local properties of the objective function
allowing accelerating the convergence essentially has been applied to the multi-objective
problems.

\keywords{deterministic global optimization, multi-objective optimization, parallel numerical
methods, derivative-free algorithms}
\end{abstract}
%
\section{Introduction}
The multi-objective optimization problems attract an increasing interest in recent years since
these ones reflect the controversy of the requirements arising to the modeled objects of real
world the most completely. In such problems, the solutions are a compromised decision of a set
of these ones. As a rule, these compromised solutions are chosen either from a set, in which all
elements cannot be improved with respect to any criterion without the worsening of the other
criteria (Pareto-optimal or efficient solutions) or from a set, in which no one element could be
improved with respect to all criteria at once (Slater-optimal or weakly efficient solutions). The
latter requirement is weaker, and all Pareto-optimal solutions are the Slater-optimal ones as
well.

To find the complete set of efficient or weakly efficient points is the most computationally
complex problem, however it provides a wide choice of possible compromised solutions. A
plenty of approaches to the finding of these sets has been developed: the application of the
genetic algorithms to the multi-objective problem directly~\cite{DebPratap2002}, the use of the
parametric convolutions reducing the problem to a series of the scalar global optimization
problems~\cite{GergelKozinov2017}, various approaches to the solving the problems with the
convex~\cite{ZhangZuo2013} or linear~\cite{Benson1998} objectives. The main problem in the
using of the most of these methods is the absence of the theoretical guarantee of the uniform
convergence to the whole sought set of the Pareto-optimal solutions on a wide class of
problems including the ones with the non-convex multiextremal criteria. A method having this
property in the case, when all criteria satisfy the Lipschitz condition has been developed on the
basis of the information global search algorithm~\cite{markinStrongin1993}. In the present
paper, the parallel and accelerated by the local refinement technique~\cite{strOptBook}(chapter
3) modification of this method will be considered.

\section{Problem Statement and Dimensionality Reduction}
The multi-objective optimization problems are stated in the following way:
\begin{equation}
  \label{eq:problem}
  \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i
\leqslant n \}.
\end{equation}
Let us assume the components of the vector function (the partial criteria) \(f_i(y), 1\leqslant
i\leqslant m\), to satisfy the Lipschitz condition with the Lipschitz constant \(L_i\) in \(D\):
\begin{displaymath}
\label{lip}
|f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,1\leqslant i
\leqslant m
\end{displaymath}

@@

\begin{equation}
  \label{eq:slater}
  S(D) = \{y\in D: \nexists z\in D, f_i(z)<f_i(y),1\leqslant i \leqslant m\},
\end{equation}
which is usually referred as the set of semi-efficient (or weakly efficient) solutions. The
conditions in the right-hand side of the definition (\ref{eq:slater}) are known as the principle of
weak Pareto-optimality (or Slater's optimality principle).

The use of the space-filling curve (\textit{evolvent}) \(y(x)\)
\begin{displaymath}
\label{cube}
\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{displaymath}
is a classic dimensionality reduction scheme for global optimization algorithms~\cite{evolvents2013}.
Such a mapping allows the reduction of the problem (\(\ref{eq:problem}\)) stated in a
multidimensional space to solving a one-dimensional problem at the expense of worsening its
properties.
In particular, the one-dimensional functions \(f_i(y(x))\) are not Lipschitzian but the
H\"olderian functions:
\begin{equation}
\label{eq:holder}
|f_i(y(x_1))-f_i(y(x_2))|\leqslant H_i{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1],
\end{equation}
where the H\"older constants \(H_i\) are related to the Lipschitz constant \(L_i\) by the relation
\begin{displaymath}
H_i=4L_id\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant n\}.
\end{displaymath}
\par
Therefore, not limiting the generality, one can consider the solving of the
one-dimensional problem \(\min\{f(y(x)): x\in [0,1]\}\), satisfying the H\"older condition. The
issues of the numerical building the mapping like a Peano curve and the corresponding theory
have been considered in detail in~\cite{evolvents2013}. Here we would note that an evolvent
built numerically is an approximation of the theoretical Peano curve with a precision of the
order \(2^{-m}\), where \(m\) is the building parameter of the evolvent.

@@

The theoretical substantiation of this method when \(p=1\) and \(q=0\) is presented in
\cite{strOptBook}(chapter 3).
%The siffitient condition of convergence is: there exists an iteration such that \(r\mu_\nu \geqslant 4H_\nu, 1\leqslant \nu \leqslant m\).
\section{Test Problems}
\label{sec:test_problems}
In order to evaluate the degree of speedup of the convergence of the modified algorithm from
Section \ref{sec:algorithm}, the following problems were used:
\begin{enumerate}
  \item Markin-Strongin problem \cite{markinStrongin1993}:
    \begin{equation}
      \left \{
      \begin{array}{l}
        f_1(y) = \min\{\sqrt{y_1^2+y_2^2},\sqrt{(y_1-1.5)^2+(y_2+1.5)^2}\} \\
        f_2(y)=\sqrt{(y_1+0.5)^2+(y_2-0.5)^2} \\
      \end{array}
      \right .
      y_1\in [-1;2],y_2\in [-2;1]
    \end{equation}
  \item Fonseca and Fleming problem \cite{Huband2006}:
  \begin{equation}
    \label{eq:fonseca}
    \left \{
    \begin{array}{l}
      f_{1}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} - \frac{1}{\sqrt{n}}
\right)^{2} \right) \\
      f_{2}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} + \frac{1}{\sqrt{n}}
\right)^{2} \right) \\
    \end{array}
    \right .
    \quad y\in [-4;4]^n
  \end{equation}
  \item
  Viennet function~\cite{Huband2006}:
  \begin{equation}
    \left \{
    \begin{array}{l}
      f_{1}\left(y\right) = 0.5(y_1^2 + y_2^2) + \sin(y_1^2 + y_2^2)\\
      f_{2}\left(y\right) = \frac{(3y_1 - 2y_2 + 4)^2}{8} + \frac{(y_1-y_2 + 1)^2}{27} + 15\\
      f_{3}\left(y\right) = \frac{1}{y_1^2+y_2^2 + 1} -1.1 \exp\{-(y_1^2+y_2^2)\}\\
    \end{array}
    \right .
    \quad y\in [-3;3]^2
  \end{equation}
%  Schaffer function N. 2 \cite{Huband2006}:
%    \begin{equation}
%      \label{eq:schaffer2}
%      \left \{
%      \begin{array}{l}
%        f_1(y) = \begin{cases}
%                  -y,   & \text{if } y \leqslant 1 \\
%                   y-2, & \text{if } 1 < y \leqslant 3 \\
%                   4-y, & \text{if } 3 < y \leqslant 4 \\
%                   y-4, & \text{if } y > 4 \\
%                \end{cases} \\
%        f_2(y) = (y-5)^2\\
%      \end{array}
%      \right .
%      \quad y\in [-5;10]
%    \end{equation}
  \item Poloni's two objective function \cite{Huband2006}:
    \begin{equation}
      \left \{
      \begin{array}{l}
        f_{1}\left(y\right) = \left[1 + \left(A_{1} - B_{1}\left(y\right) \right)^{2} + \left(A_{2} -
B_{2}\left(y\right) \right)^{2} \right] \\
        f_{2}\left(y\right) = \left(y_1 + 3\right)^{2} + \left(y_2 + 1 \right)^{2} \\
      \end{array}
      \right .
      \quad y\in [-\pi;\pi]^2
    \end{equation}
    where
    \begin{equation*}
      \begin{cases}
        A_{1} = 0.5 \sin \left(1\right) - 2 \cos \left(1\right) + \sin \left(2\right) - 1.5 \cos
\left(2\right)  \\
        A_{2} = 1.5 \sin \left(1\right) - \cos \left(1\right) + 2 \sin \left(2\right) - 0.5 \cos
\left(2\right)  \\
        B_{1}\left(y\right) = 0.5 \sin \left(y_1\right) - 2 \cos \left(y_1\right) + \sin \left(y_2\right) -
1.5 \cos \left(y_2\right)  \\
        B_{2}\left(y\right) = 1.5 \sin \left(y_1\right) - \cos \left(y_1\right) + 2 \sin \left(y_2\right) -
0.5 \cos \left(y_2\right)
      \end{cases}
    \end{equation*}
%  \item A problem considered in \cite{}:
%    \begin{equation}
%      \left \{
%      \begin{array}{l}
%        f_1(y) = (y_1 - 1)y_2^2 +1 \\
%        f_2(y)= y_2 \\
%      \end{array}
%      \right .
%      \quad y\in [0,1]^2
%    \end{equation}
\end{enumerate}
\section{Experimental Results}
The computational experiments have been carried out on the Lobachevsky supercomputer at
State University of Nizhny Novgorod. A computational node included 2 Intel
Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores
were available per a node).
In the present section, we will understand the speedup of the method as the speedup in the
number of executed iterations, not in the time of execution. If the problem
criteria (\ref{eq:problem}) is time-consuming, the overhead costs for the execution of the
decision rules of the optimization method are low as compared to the time of computing the
criteria \(f_i(y), 1\leqslant i\leqslant m\).

\paragraph{Local refinement advantages.} In \cite{barkalovLebedev2016} the numerical
experiments demonstrating the speedup of convergence when using the local refinement
technique in the algorithm solving the scalar global optimization problems are presented. One
could expect similar results from the multi-objective algorithm as well. The multi-objective
algorithm with local refinement (MOAR) has been applied to the  Fonseca and Fleming
problem (\ref{eq:fonseca}) at \(n=2\). The parameters of the method were as follows:
\(\varepsilon=0.01,\;r=4,\;q=4,\;\alpha=15,\;p=1\). Before the method stops, 1176 iterations
have been executed, the number of found weakly optimal points was 90. At \(q=0\) (without the
local refinement) the method had performed 1484 iterations, the number of found weakly
optimal points was 93. In Fig. \ref{fig:fonseca_slater} the set \(S\) in the problem
(\ref{eq:fonseca}) and the numerical estimates of this one obtained at \(q=0\) (Fig.
\ref{fig:fonseca_slater}.a) and \(q=4\) (Fig. \ref{fig:fonseca_slater}.b) are presented. It is
evident also from Fig. \ref{fig:fonseca_slater} that the approximate solution covers the whole
set of the weak-optimal solutions.

@@

\paragraph{Parallel method results.} In order to demonstrate the speedup in iterations, which
MOAR provides when \(p > 1\), all problems from Section \ref{sec:test_problems} have been
solved for the values \(p=1,2,4,8,16\). The parameters of the method were the following:
\(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). The number of iterations is presented in Table \ref{tab:exResultsIters}
(in the braces, the cardinality of the set \(S\) is given) and the speedup in iterations is presented
in in Table \ref{tab:exResultsItersSpeedup}. As it is evident form Table
\ref{tab:exResultsIters}, the cardinality of the weak-optimal solutions set changed insufficiently when varying \(p\)
i. e. the quality of the obtained estimates remained the same. At the same time, the number of
iterations decreased linearly with increasing value of \(p\) (Table
\ref{tab:exResultsItersSpeedup}). In Fig. \ref{fig:s_estimation_all}, the examples of the
numerical solutions of the considered problems are given.

@@

\section{Conclusion}
In the present work, a method for solving the multi-objective optimization problems allowing
finding a uniform estimate of the set of the weakly-optimal points has been considered. The
techniques of parallelization and of the acceleration of convergence common for all algorithms
of similar structure have been applied. The numerical experiments performed have
demonstrated a speedup in convergence when using the local refinement. Also, the
parallelization based on the characteristics was found to be efficient for this method. The
speedup in iterations at the parallelization based on the characteristics appeared to be linear in
most cases as for the information methods of scalar optimization (see the results of the
parallelization of the scalar optimization method, for example, in
\cite{barkalovLebedef2016}). From the results of the experiments, one can conclude that it is
expedient to apply the parallel MOAR in the problems of low dimensionality (1 to 3) with
the computation-costly criteria. Moreover, the property of uniform convergence of the considered
algorithm to the whole set of the weakly-optimal solutions has been demonstrated on the
numerical examples.

>>>>

\begin{abstract}
This work considers a parallel algorithm for the multi-objective optimization. The
algorithm is based on the application of the information-statistical approach for solving
reduced scalar problem in which the set of the global optima coincides to the set of the
weakly efficient solutions of the initial multi-objective problem.
In the work the parallelization scheme, which is common for all information-statistical global search algorithms,
was applied to the sequential algorithm of the multi-objective optimization. Also, in this
work one of the convergence speedup techniques based on local properties of objective functions has been applied to the multi-objective
problems.

\keywords{deterministic global optimization, multi-objective optimization, parallel numerical
methods, derivative-free algorithms}
\end{abstract}
%
\section{Introduction}
The multi-objective optimization problems attract an increasing interest in recent years since
these ones most completely reflect the controversy of the requirements arising to the modeled objects of real
world. In such problems, the solutions are a compromised decision of a set
of these ones. As a rule, these compromised solutions are chosen either from a set, in which all
elements cannot be improved with respect to any criterion without the worsening of the other
criteria (Pareto-optimal or efficient solutions) or from a set, in which no one element could be
improved with respect to all criteria at once (Slater-optimal or weakly efficient solutions). The
latter requirement is weaker, and all Pareto-optimal solutions are the Slater-optimal ones as
well.

To find the complete set of efficient or weakly efficient points is the most computationally
complex problem, however it provides a wide choice of possible compromised solutions. To find these sets a
plenty of approaches have been developed: the application of the
genetic algorithms to the multi-objective problem directly~\cite{DebPratap2002}, the use of the
parametric convolutions reducing the problem to a series of the scalar global optimization
problems~\cite{GergelKozinov2017}, various approaches to the solving the problems with the
convex~\cite{ZhangZuo2013} or linear~\cite{Benson1998} objectives. The main problem in the
using of the most of these methods is the absence of the theoretical guarantee of the uniform
convergence to the whole sought set of the Pareto-optimal solutions on a wide class of
problems including the ones with the non-convex multiextremal criteria. A method having this
property in the case, when all criteria satisfy the Lipschitz condition has been developed on the
basis of the information global search algorithm~\cite{markinStrongin1993}.
In the paper there is considered the parallel modification of this method accelerated due to the local refinement
technique~\cite{strOptBook}(chapter 3).

\section{Problem Statement and Dimensionality Reduction}
The multi-objective optimization problems are stated in the following way:
\begin{equation}
  \label{eq:problem}
  \min\{f(y): y\in D\}, D=\{y\in \mathbb{R}^n: a_i \leqslant y_i \leqslant b_i, 1\leqslant i
\leqslant n \}.
\end{equation}
Assume that the components of the vector function (the partial criteria) \(f_i(y), 1\leqslant
i\leqslant m\) satisfy the Lipschitz condition with the Lipschitz constant \(L_i\) in \(D\):
\begin{displaymath}
\label{lip}
|f_i(y_1)-f_i(y_2)|\leqslant L_i\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L_i<\infty,1\leqslant i
\leqslant m
\end{displaymath}

@@

\begin{equation}
  \label{eq:slater}
  S(D) = \{y\in D: \nexists z\in D, f_i(z)<f_i(y),1\leqslant i \leqslant m\},
\end{equation}
which is called the set of semi-efficient (or weakly efficient) solutions. The
conditions in the right-hand side of the definition (\ref{eq:slater}) are known as the principle of
weak Pareto-optimality (or Slater's optimality principle).

The use of the space-filling curve (\textit{evolvent}) \(y(x)\)
\begin{displaymath}
\label{cube}
\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant
N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{displaymath}
is a classic dimensionality reduction scheme for global optimization algorithms~\cite{evolvents2013}.
Such a mapping allows to reduce the problem (\(\ref{eq:problem}\)) stated in a
multidimensional space to solving a one-dimensional problem at the expense of worsening its
properties.
In particular, the one-dimensional functions \(f_i(y(x))\) are not Lipschitzian but the
H\"olderian functions:
\begin{equation}
\label{eq:holder}
|f_i(y(x_1))-f_i(y(x_2))|\leqslant H_i{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1],
\end{equation}
where the H\"older constants \(H_i\) are related to the Lipschitz constant \(L_i\) by the relation
\begin{displaymath}
H_i=4L_id\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant n\}.
\end{displaymath}
\par
Without losing generality one can, one can consider the solving of the
one-dimensional problem \(\min\{f(y(x)): x\in [0,1]\}\), satisfying the H\"older condition. The
issues of the numerical building the mapping like a Peano curve and the corresponding theory
have been considered in detail in~\cite{evolvents2013}. Here we would note that an evolvent
built numerically approximates the theoretical Peano curve with a precision of the
order \(2^{-m}\), where \(m\) is the building parameter of the evolvent.

@@

The theoretical substantiation of this method when \(p=1\) and \(q=0\) is presented in
\cite{strOptBook}(chapter 3).
%The siffitient condition of convergence is: there exists an iteration such that \(r\mu_\nu \geqslant 4H_\nu, 1\leqslant \nu \leqslant m\).
\section{Test Problems}
\label{sec:test_problems}
To evaluate the degree of speedup of the convergence of the modified algorithm from
Section \ref{sec:algorithm}, the following problems were used:
\begin{enumerate}
  \item Markin-Strongin problem \cite{markinStrongin1993}:
    \begin{equation}
      \left \{
      \begin{array}{l}
        f_1(y) = \min\{\sqrt{y_1^2+y_2^2},\sqrt{(y_1-1.5)^2+(y_2+1.5)^2}\} \\
        f_2(y)=\sqrt{(y_1+0.5)^2+(y_2-0.5)^2} \\
      \end{array}
      \right .
      y_1\in [-1;2],y_2\in [-2;1]
    \end{equation}
  \item Fonseca and Fleming problem \cite{Huband2006}:
  \begin{equation}
    \label{eq:fonseca}
    \left \{
    \begin{array}{l}
      f_{1}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} - \frac{1}{\sqrt{n}}
\right)^{2} \right) \\
      f_{2}\left(y\right) = 1 - \exp \left(-\sum_{i=1}^{n} \left(y_{i} + \frac{1}{\sqrt{n}}
\right)^{2} \right) \\
    \end{array}
    \right .
    \quad y\in [-4;4]^n
  \end{equation}
  \item
  Viennet function~\cite{Huband2006}:
  \begin{equation}
    \left \{
    \begin{array}{l}
      f_{1}\left(y\right) = 0.5(y_1^2 + y_2^2) + \sin(y_1^2 + y_2^2)\\
      f_{2}\left(y\right) = \frac{(3y_1 - 2y_2 + 4)^2}{8} + \frac{(y_1-y_2 + 1)^2}{27} + 15\\
      f_{3}\left(y\right) = \frac{1}{y_1^2+y_2^2 + 1} -1.1 \exp\{-(y_1^2+y_2^2)\}\\
    \end{array}
    \right .
    \quad y\in [-3;3]^2
  \end{equation}
%  Schaffer function N. 2 \cite{Huband2006}:
%    \begin{equation}
%      \label{eq:schaffer2}
%      \left \{
%      \begin{array}{l}
%        f_1(y) = \begin{cases}
%                  -y,   & \text{if } y \leqslant 1 \\
%                   y-2, & \text{if } 1 < y \leqslant 3 \\
%                   4-y, & \text{if } 3 < y \leqslant 4 \\
%                   y-4, & \text{if } y > 4 \\
%                \end{cases} \\
%        f_2(y) = (y-5)^2\\
%      \end{array}
%      \right .
%      \quad y\in [-5;10]
%    \end{equation}
  \item Poloni's two objective function \cite{Huband2006}:
    \begin{equation}
      \left \{
      \begin{array}{l}
        f_{1}\left(y\right) = \left[1 + \left(A_{1} - B_{1}\left(y\right) \right)^{2} + \left(A_{2} -
B_{2}\left(y\right) \right)^{2} \right] \\
        f_{2}\left(y\right) = \left(y_1 + 3\right)^{2} + \left(y_2 + 1 \right)^{2} \\
      \end{array}
      \right .
      \quad y\in [-\pi;\pi]^2
    \end{equation}
    where
    \begin{equation*}
      \begin{cases}
        A_{1} = 0.5 \sin \left(1\right) - 2 \cos \left(1\right) + \sin \left(2\right) - 1.5 \cos
\left(2\right)  \\
        A_{2} = 1.5 \sin \left(1\right) - \cos \left(1\right) + 2 \sin \left(2\right) - 0.5 \cos
\left(2\right)  \\
        B_{1}\left(y\right) = 0.5 \sin \left(y_1\right) - 2 \cos \left(y_1\right) + \sin \left(y_2\right) -
1.5 \cos \left(y_2\right)  \\
        B_{2}\left(y\right) = 1.5 \sin \left(y_1\right) - \cos \left(y_1\right) + 2 \sin \left(y_2\right) -
0.5 \cos \left(y_2\right)
      \end{cases}
    \end{equation*}
%  \item A problem considered in \cite{}:
%    \begin{equation}
%      \left \{
%      \begin{array}{l}
%        f_1(y) = (y_1 - 1)y_2^2 +1 \\
%        f_2(y)= y_2 \\
%      \end{array}
%      \right .
%      \quad y\in [0,1]^2
%    \end{equation}
\end{enumerate}
\section{Experimental Results}
The computational experiments have been carried out on the Lobachevsky supercomputer at
State University of Nizhny Novgorod. A computational node included 2 Intel
Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores
were available per a node). C++ language and OpenMP parallel framework were used to implement the considered algorithm.

In this section, we will understand the speedup of the method as the speedup in the
number of executed iterations, not in the time of execution. If the problem
criteria (\ref{eq:problem}) is time-consuming, the overhead costs for the execution of the
decision rules of the optimization method are low as compared to the time of computing the
criteria \(f_i(y), 1\leqslant i\leqslant m\).

\paragraph{Local refinement advantages.} In \cite{barkalovLebedev2016} the numerical experiments
are presented which demonstrate the speedup of convergence of the scalar global optimization algorithm using the local refinement techniques described above.
One could expect similar results from the multi-objective algorithm as well. The multi-objective
algorithm with local refinement (MOAR) has been applied to the  Fonseca and Fleming
problem (\ref{eq:fonseca}) at \(n=2\). The parameters of the method were as follows:
\(\varepsilon=0.01,\;r=4,\;q=4,\;\alpha=15,\;p=1\). Before the method stops, 1176 iterations
have been executed, the number of found weakly optimal points was 90. At \(q=0\) (without the
local refinement) the method had performed 1484 iterations, the number of found weakly
optimal points was 93. In Fig. \ref{fig:fonseca_slater} the set \(S\) in the problem
(\ref{eq:fonseca}) and the numerical estimates of this one obtained at \(q=0\) (Fig.
\ref{fig:fonseca_slater}.a) and \(q=4\) (Fig. \ref{fig:fonseca_slater}.b) are presented. It is
evident also from Fig. \ref{fig:fonseca_slater} that the approximate solution covers the whole
set of the weak-optimal solutions.

@@

\paragraph{Parallel method results.} In order to demonstrate the speedup in iterations, which
MOAR provides when \(p > 1\), all problems from Section \ref{sec:test_problems} have been
solved for the values \(p=1,2,4,8,16\). The parameters of the method were the following:
\(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). The number of iterations is presented in Table \ref{tab:exResultsIters}
(in the braces, the cardinality of the set \(S\) is given) and the speedup in iterations is presented
in in Table \ref{tab:exResultsItersSpeedup}. As it is evident form Table
\ref{tab:exResultsIters}, the cardinality of the weak-optimal solutions set changed insufficiently when varying \(p\)
i. e. the quality of the obtained estimates remained the same. At the same time, the number of
iterations decreased linearly with increasing value of \(p\) (Table
\ref{tab:exResultsItersSpeedup}). Thus, the parallel algorithm reduces the number of iterations almost in \(p\) times comparing to the sequantal one. In Fig. \ref{fig:s_estimation_all}, the examples of the
numerical solutions of the considered problems are given.

@@

\section{Conclusion}
In the work a method for solving the multi-objective optimization problem has been considered which allows to
find a uniform estimate of the set of the weakly-optimal points.
The techniques of parallelization and of the acceleration of convergence common for all algorithms
of similar structure have been applied. The numerical experiments performed have
demonstrated a speedup in convergence when using the local refinement. Also, the
parallelization based on the characteristics was found to be efficient for this method. The
speedup in iterations at the parallelization based on the characteristics appeared to be linear in
most cases as for the information methods of scalar optimization (see the results of the
parallelization of the scalar optimization method, for example, in
\cite{barkalovLebedef2016}). From the results of the experiments, one can conclude that it is
expedient to apply the parallel MOAR in the problems of low dimensionality (1 to 3) with
the computation-costly criteria. Moreover, the property of uniform convergence of the considered
algorithm to the whole set of the weakly-optimal solutions has been demonstrated on the
numerical examples.
